### 2021-11-10 Rmarkdown/preprocessing/Master_Phenotype_File_211108.Rmd
Update microbiome workflow.

---
title: "Master Pheno FIle"
author: "Kristen James"
date: "11/8/2021"
output: html_document
---

```{r Load TMAO pheno file}
tmao <- read.csv("../../data/raw/from_RedCAP/TMAO/CTSC24532USDAWHNRCNu-TMAOKJ012220_DATA_2020-07-15_0915_NAedited.csv", row.names = 1)
```

Load tobacco data and combine it with the master object.
```{r Load tobbaco data, results='hide'}
dem <- read.csv("../../data/raw/from_RedCAP/demographic/CTSC24532USDAWHNRCNu-DemographicData_DATA_2020-08-28_1058.csv")

dem_tobbaco <- dem[,c(1,26)] #26 in tobbaco use
rownames(dem_tobbaco) <- dem_tobbaco[,"subject_id"]
dem_tobbaco <- subset(dem_tobbaco, select = -c(subject_id))

master <- merge(tmao, dem_tobbaco, by = 0)
rownames(master) <- master[,1]
master <- subset(master, select=-c(Row.names))
```

Load ethnicity data and combine it with the master object.
```{r Load ethnicity data, results='hide'}
eth <- read.csv("../../data/processed/ethnicity/ethnicity_210715.csv",row.names = 1)

# We will reorder the levels so that it matches the order from the demographic survey. This may help prevent mistakes in the future.
str(eth$ethnicity)
eth$ethnicity <- ordered(eth$ethnicity, levels = c("Caucasian", "Hispanic", "African.American", "Asian", "Multi", "Other", "NA"))
# This also turns ethnicity into a factor (not a character)

# Merge
master <- merge(master, eth, by = 0)
rownames(master) <- master[,1]
master <- subset(master, select=-c(Row.names))
```

```{r}
colnames(master)
```
Save
```{r}
write.csv(master, "../../data/processed/phenotype/Phenotypes_211108.csv")
saveRDS(master, "../../data/processed/phenotype/Phenotypes_211108.rds")
```