### 2021-11-12 Rmarkdown/analysis/Characteristics_Diet_CardiometabolicRisk_211111.Rmd
Update analysis workflows

---
title: "Describe Participants: Characteristics, Diet, and Cardiometabolic Risk"
author: "Kristen James"
date: "11/11/2021"
output: html_document
---
# About

**Project:** FL100

**About:** Objectives of the paper analyzed in this markdown include:

1) Who are the participants in the study? --> Demonstrate who is in the study by making tables and plots for the manuscript.
2) What do they eat? --> Describe food intake by TMAO quartile. Complete multiple linear regression analysis.
3) Are TMAO levels related to cardiometabolic risk? --> Complete multiple linear regression analysis.

**Inputs:** Inputs include:

1) Master phenotype data
2) ASA24 averaged data with transformations

**Outputs:** Outputs include:

1)

**Sources:**

```{r Libraries, warning=FALSE, message=FALSE}
library(plyr)
library(tableone)
```

```{r Load data}
phen <- readRDS("../../data/processed/phenotype/Phenotypes_211108.rds")

# Complete TMAO data only
phen <- phen[complete.cases(phen$tmao == TRUE),]

# Get rid of NA ethnicity category
phen$ethnicity <- droplevels(phen$ethnicity)

# Factor and recode sex
phen$sex <- as.factor(phen$sex)
phen$sex <- plyr::revalue(phen$sex, c("1" = "Male", "2" = "Female"))

# Make TMAO log variable
phen$tmao_log <- log(phen$tmao)
```

# Descriptive Characteristics

## Table 1. Descriptive characteristics of participants

Note, crp is in units of ng/mL. Report in mg/dL so multiply by 0.0001.
il6 is in units of pg/mL
tnfa is in units of pg/mL

```{r Convert crp units}
phen$crp_bd1.mgdL <- phen$crp_bd1 * 0.0001
```

```{r Participant characterisitics}
# TMAO table by age bin - females
factorVars <- c("age_cat")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al", "hdl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.f <- CreateTableOne(vars = vars,
                                    strata = "age_cat",
                                    data = phen[phen$sex == "Female",],
                                    factorVars = factorVars)
table_tmao.f

# TMAO table by age bin - males
factorVars <- c("age_cat")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al", "hdl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.m <- CreateTableOne(vars = vars,
                                    strata = "age_cat",
                                    data = phen[phen$sex == "Male",],
                                    factorVars = factorVars)
table_tmao.m

# TMAO table by sex
factorVars <- c("sex")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al", "hdl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.all <- CreateTableOne(vars = vars,
                                    strata = "sex",
                                    data = phen,
                                    factorVars = factorVars)
table_tmao.all
```

Save the tables

```{r Save characteristic tables}
table_save.f <- print(table_tmao.f, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.f, file = "../../Outputs/tables/ParticipantChar/Characteristics_Female_byAgeBin_211111.csv")

table_save.m <- print(table_tmao.m, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.m, file = "../../Outputs/tables/ParticipantChar/Characteristics_Male_byAgeBin_211111.csv")

table_save.all <- print(table_tmao.all, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.all, file = "../../Outputs/tables/ParticipantChar/Characteristics_All_bySex_211111.csv")
```

Is TMAO significantly different by ethnicity?
```{r Ethnicity}
ethnicity_aov <- (aov(tmao_log ~ ethnicity, phen))
summary(ethnicity_aov)
TukeyHSD(ethnicity_aov)

ethnicity_lm <- (lm(tmao_log ~ ethnicity + sex:age + cystatinc_bd1, phen))
summary(ethnicity_lm)
```

There are significant differences in TMAO levels by ethnic groups but ethnicity does not predict TMAO.

Back to tables... make similar tables but show TMAO quantiles

```{r Quantile TMAO}
tmao_quantile <- quantile(phen$tmao)
phen$tmao_quantile <- ifelse(phen$tmao <= tmao_quantile[2], "Quantile1",
                             ifelse(phen$tmao > tmao_quantile[2] & phen$tmao <= tmao_quantile[3], "Quantile2",
                                   ifelse(phen$tmao > tmao_quantile[3] & phen$tmao <= tmao_quantile[4], "Quantile3", "Quantile4")))

# Check it
table(phen$tmao_quantile)
phen$tmao_quantile <- as.factor(phen$tmao_quantile)
str(phen$tmao_quantile)
```

```{r Participant characterisitics by TMAO quantile}
# TMAO table by age bin - females
factorVars <- c("age_cat", "tmao_quantile")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al", "hdl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.fq <- CreateTableOne(vars = vars,
                                    strata = "tmao_quantile",
                                    data = phen[phen$sex == "Female",],
                                    factorVars = factorVars)
table_tmao.fq

# TMAO table by age bin - males
factorVars <- c("age_cat", "tmao_quantile")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al", "hdl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.mq <- CreateTableOne(vars = vars,
                                    strata = "tmao_quantile",
                                    data = phen[phen$sex == "Male",],
                                    factorVars = factorVars)
table_tmao.mq

# TMAO table by sex
factorVars <- c("tmao_quantile")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al", "hdl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.allq <- CreateTableOne(vars = vars,
                                    strata = "tmao_quantile",
                                    data = phen,
                                    factorVars = factorVars)
table_tmao.allq
```

```{r Save characteristic quantile tables}
table_save.fq <- print(table_tmao.fq, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.fq, file = "../../Outputs/tables/ParticipantChar/Characteristics_Female_byTMAOQuantile_211111.csv")

table_save.mq <- print(table_tmao.mq, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.mq, file = "../../Outputs/tables/ParticipantChar/Characteristics_Male_byTMAOQuantile_211111.csv")

table_save.allq <- print(table_tmao.allq, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.allq, file = "../../Outputs/tables/ParticipantChar/Characteristics_All_byTMAOQuantile_211111.csv")
```

# Cardiometabolic Risk Analysis

```{r Cardiometabolic Multiple Linear Regression}
# Checked all variables for normality via shapiro.test() in console. TG, and glc ere the only ones that didn't pass.
phen$glc_bd1_ln <- log(phen$glc_bd1)
phen$tg_bd1_ln <- log(phen$tg_bd1)
phen$crp_bd1_ln <- log(phen$crp_bd1)
phen$tnfa_bd1_ln <- log(phen$tnfa_bd1)
phen$il6_bd1_ln <- log(phen$il6_bd1)

transformed_keep_cardio_tbl <- c("sex", "age","ethnicity", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al75", "hdl_bd1", "ldl_bd1", "chol_bd1", "tg_bd1_ln", "glc_bd1_ln","tnfa_bd1_ln", "il6_bd1_ln", "crp_bd1_ln", "cystatinc_bd1", "choline", "betaine", "carnitine","tmao_log")

#~~~~~~~~
# TMAO  
#~~~~~~~~
# Make a smaller df from master (the old data)
phen_sub <- phen[, colnames(phen) %in% transformed_keep_cardio_tbl]
cvd_biomark <- subset(phen_sub, select = -c(sex, age, ethnicity, cystatinc_bd1, tmao_log)) # change variable here
n <- ncol(cvd_biomark)

# LCMS method
# Run linear model
my_lm <- lapply(1:n, function(x) lm(cvd_biomark[,x] ~ tmao_log + sex*age + cystatinc_bd1 + ethnicity, phen_sub )) # include kcal

# Organize results
my_lm2 <- sapply(my_lm, function(x){summary(x)$adj.r.squared})
my_lm2.5 <- sapply(my_lm, function(x){summary(x)$coefficients[2,1]}) # [CVD row, estimate] corresponds to CVD B 
my_lm2.6 <- sapply(my_lm, function(x){summary(x)$coefficients[2,2]}) # [CVD row, Std.Error] corresponds to CVD Std. Error 
my_lm2.7 <- sapply(my_lm, function(x){summary(x)$coefficients[2,3]}) # [CVD row, T value] corresponds to CVD t value
my_lm3 <- sapply(my_lm, function(x){summary(x)$coefficients[2,4]}) # [CVD row, P] corresponds to CVD P value
my_lm361_lcms_cvd <- cbind(my_lm2, my_lm2.5, my_lm2.6, my_lm2.7, my_lm3)
colnames(my_lm361_lcms_cvd) <- c("adj.r.squared", "estimate", "Std.Error", "T.value","P")
row.names(my_lm361_lcms_cvd) <- colnames(cvd_biomark)
res <- as.data.frame(my_lm361_lcms_cvd)

# BH correction
res$p.adj <- stats::p.adjust(res$P, method = "BH", n = n)

# Save table
write.csv(res, "../../Outputs/tables/CardiometabolicMarkers/MultipleLinReg_TMAO_LCMS_n361_CVD_211111.csv")
```

# Diet Analysis

Load the diet data.

```{r Load ASA24 averaged data}
food <- readRDS(file = "../../data/processed/asa24/ASA24_meanIntake_withTransformations_211111.rds")

# Combine food with phenotype data
full <- merge(phen, food, by = 0)
row.names(full) <- full[,"Row.names"]
full <- subset(full, select = -c(Row.names))
```

Make a list of foods and covariate variables to keep.

```{r Foods to keep, results='hide'}
transformed_keep_tbl <- c("sex", 
#"sex_numeric",
"kcal_tnfs", 
"age", 
"cystatinc_bd1", 
"ethnicity",
"tmao", 
"tmao_log", 
#"TMAO.0",
#"TMAO.30",
#"TMAO.180",
#"TMAO.360",
#"lnTMAO.0",
#"lnTMAO.30",
#"lnTMAO.180",
#"lnTMAO.360",
#"TMAO_0_30",
#"TMAO_0_180",
#"TMAO_0_360",
#"Ind.SO4",
#"p.Cresol.SO4",
#"TMAO_quart_num",
#"T.pf_mps_totals_sqrRt",
"T.pf_meat_totals_sqrRt",
"T.pf_curedmeat_totals_sqrRt",
"T.pf_poult_totals_sqrRt",
"T.pf_seafd_hi_totals_Binary",
"T.pf_seafd_low_totals_Binary",
"T.pf_eggs_totals_sqrRt",
"T.pf_nutsds_totals_sqrRt",
"T.pf_legumes_totals_sqrRt",
"T.pf_total_totals_natLog",
#"T.pf_soy_totals_Binary",
"T.v_drkgr_totals_sqrRt",
"T.v_redor_total_totals_sqrRt",
#"T.v_redor_tomato_totals_sqrRt",
#"T.v_redor_other_totals_sqrRt",
"T.v_starchy_total_totals_sqrRt",
#"T.v_starchy_potato_totals_sqrRt",
#"T.v_starchy_other_totals_sqrRt",
"T.v_other_totals_sqrRt",
"T.v_total_totals_sqrRt",
#"T.v_legumes_total_sqrRt",
"T.g_whole_totals_sqrRt",
"T.g_refined_totals_sqrRt",
"g_total_totals",
"T.d_total_totals_sqrRt",
"T.d_milk_totals_sqrRt",
"T.d_cheese_totals_sqrRt",
"T.d_yogurt_totals_Binary"
)
```

Now run the multiple linear regression.

```{r TMAO ~ food - LCMS BD1 - lm 356 ppl, results='hide'}
# Make a smaller df from master (the old data)
full_sub <- full[, colnames(full) %in% transformed_keep_tbl]
food <- subset(full_sub, select = -c(sex, kcal_tnfs, age, cystatinc_bd1, ethnicity, tmao, tmao_log))
n <- ncol(food)

# LCMS method
# Run linear model
my_lm <- lapply(1:n, function(x) lm(tmao_log ~ food[,x] + sex*age + cystatinc_bd1 + ethnicity + kcal_tnfs, full_sub )) # include kcal

# Organize results
my_lm2 <- sapply(my_lm, function(x){summary(x)$adj.r.squared})
my_lm2.5 <- sapply(my_lm, function(x){summary(x)$coefficients[2,1]}) # [food row, estimate] corresponds to food B 
my_lm2.6 <- sapply(my_lm, function(x){summary(x)$coefficients[2,2]}) # [food row, Std.Error] corresponds to food Std. Error 
my_lm2.7 <- sapply(my_lm, function(x){summary(x)$coefficients[2,3]}) # [food row, T value] corresponds to food t value
my_lm3 <- sapply(my_lm, function(x){summary(x)$coefficients[2,4]}) # [food row, P] corresponds to food P value
my_lm356_lcms <- cbind(my_lm2, my_lm2.5, my_lm2.6, my_lm2.7, my_lm3)
colnames(my_lm356_lcms) <- c("adj.r.squared", "estimate", "Std.Error", "T.value","P")
row.names(my_lm356_lcms) <- colnames(food)
res_food <- as.data.frame(my_lm356_lcms)

# BH correction
res_food$p.adj <- stats::p.adjust(res_food$P, method = "BH", n = n)

# Save table
write.csv(res_food, "../../Outputs/tables/Diet/MultipleLinReg_TMAO_LCMS_n356_ASA24_211111.csv")
```