### 2021-12-01 Rmarkdown/analysis/Characteristics_Diet_CardiometabolicRisk_211111.Rmd
Add dietary choline to analysis

---
title: "Describe Participants: Characteristics, Diet, and Cardiometabolic Risk"
author: "Kristen James"
date: "11/11/2021"
output: html_document
---
# About

**Project:** FL100

**About:** Objectives of the paper analyzed in this markdown include:

1) Who are the participants in the study? --> Demonstrate who is in the study by making tables and plots for the manuscript.
2) What do they eat? --> Describe food intake by TMAO quartile. Complete multiple linear regression analysis.
3) Are TMAO levels related to cardiometabolic risk? --> Complete multiple linear regression analysis.

**Inputs:** Inputs include:

1) Master phenotype data
2) ASA24 averaged data with transformations

**Outputs:** Outputs include:

1)

**Sources:**

```{r Libraries, warning=FALSE, message=FALSE}
library(plyr)
library(tableone)
```

```{r Load data}
phen <- readRDS("../../data/processed/phenotype/Phenotypes_211108.rds")

# Complete TMAO data only
phen <- phen[complete.cases(phen$tmao == TRUE),]

# Get rid of NA ethnicity category
phen$ethnicity <- droplevels(phen$ethnicity)

# Factor and recode sex
phen$sex <- as.factor(phen$sex)
phen$sex <- plyr::revalue(phen$sex, c("1" = "Male", "2" = "Female"))

# Make TMAO log variable
phen$tmao_log <- log(phen$tmao)

```

# Descriptive Characteristics

## Table 1. Descriptive characteristics of participants

Note, crp is in units of ng/mL. Report in mg/dL so multiply by 0.0001.
il6 is in units of pg/mL
tnfa is in units of pg/mL

```{r Convert crp units}
phen$crp_bd1.mgdL <- phen$crp_bd1 * 0.0001
```

```{r Participant characterisitics}
# TMAO table by age bin - females
factorVars <- c("age_cat")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al", "hdl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.f <- CreateTableOne(vars = vars,
                                    strata = "age_cat",
                                    data = phen[phen$sex == "Female",],
                                    factorVars = factorVars)
table_tmao.f

# TMAO table by age bin - males
factorVars <- c("age_cat")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al", "hdl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.m <- CreateTableOne(vars = vars,
                                    strata = "age_cat",
                                    data = phen[phen$sex == "Male",],
                                    factorVars = factorVars)
table_tmao.m

# TMAO table by sex
factorVars <- c("sex")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al", "hdl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.all <- CreateTableOne(vars = vars,
                                    strata = "sex",
                                    data = phen,
                                    factorVars = factorVars)
table_tmao.all
```

Save the tables

```{r Save characteristic tables}
table_save.f <- print(table_tmao.f, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.f, file = "../../Outputs/tables/ParticipantChar/Characteristics_Female_byAgeBin_211111.csv")

table_save.m <- print(table_tmao.m, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.m, file = "../../Outputs/tables/ParticipantChar/Characteristics_Male_byAgeBin_211111.csv")

table_save.all <- print(table_tmao.all, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.all, file = "../../Outputs/tables/ParticipantChar/Characteristics_All_bySex_211111.csv")
```

Is TMAO significantly different by ethnicity?
```{r Ethnicity}
ethnicity_aov <- (aov(tmao_log ~ ethnicity, phen))
summary(ethnicity_aov)
TukeyHSD(ethnicity_aov)

ethnicity_lm <- (lm(tmao_log ~ ethnicity + sex*age + cystatinc_bd1, phen))
summary(ethnicity_lm)
```

There are significant differences in TMAO levels by ethnic groups but ethnicity does not predict TMAO.

Back to tables... make similar tables but show TMAO quantiles

```{r Quantile TMAO}
# Quantile
tmao_quantile <- quantile(phen$tmao)
phen$tmao_quantile <- ifelse(phen$tmao <= tmao_quantile[2], "Quantile1",
                             ifelse(phen$tmao > tmao_quantile[2] & phen$tmao <= tmao_quantile[3], "Quantile2",
                                   ifelse(phen$tmao > tmao_quantile[3] & phen$tmao <= tmao_quantile[4], "Quantile3", "Quantile4")))

# Check it
table(phen$tmao_quantile)
phen$tmao_quantile <- as.factor(phen$tmao_quantile)
str(phen$tmao_quantile)

# Tertile
tmao_tertile <- quantile(phen$tmao, c(0:3/3))
phen$tmao_tertile <- ifelse(phen$tmao <= tmao_tertile[2], "Tertile1",
                             ifelse(phen$tmao > tmao_tertile[2] & phen$tmao <= tmao_tertile[3], "Tertile2", "Tertile3"))
# Factor it
phen$tmao_tertile <- factor(phen$tmao_tertile)
table(phen$tmao_tertile)

# Print tertiles for manuscript
tmao_tertile

# Checked all variables for normality via shapiro.test() in console. TG, and glc ere the only ones that didn't pass.
phen$glc_bd1_ln <- log(phen$glc_bd1)
phen$tg_bd1_ln <- log(phen$tg_bd1)
phen$crp_bd1_ln <- log(phen$crp_bd1)
phen$tnfa_bd1_ln <- log(phen$tnfa_bd1)
phen$il6_bd1_ln <- log(phen$il6_bd1)
```

```{r Participant characterisitics by TMAO quantile}
# TMAO table by age bin - females
factorVars <- c("age_cat", "tmao_quantile")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al", "hdl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.fq <- CreateTableOne(vars = vars,
                                    strata = "tmao_quantile",
                                    data = phen[phen$sex == "Female",],
                                    factorVars = factorVars)
table_tmao.fq

# TMAO table by age bin - males
factorVars <- c("age_cat", "tmao_quantile")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al", "hdl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.mq <- CreateTableOne(vars = vars,
                                    strata = "tmao_quantile",
                                    data = phen[phen$sex == "Male",],
                                    factorVars = factorVars)
table_tmao.mq

# TMAO table by sex
factorVars <- c("tmao_quantile")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al", "hdl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.allq <- CreateTableOne(vars = vars,
                                    strata = "tmao_quantile",
                                    data = phen,
                                    factorVars = factorVars)
table_tmao.allq
```

Save

```{r Save characteristic quantile tables}
table_save.fq <- print(table_tmao.fq, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.fq, file = "../../Outputs/tables/ParticipantChar/Characteristics_Female_byTMAOQuantile_211111.csv")

table_save.mq <- print(table_tmao.mq, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.mq, file = "../../Outputs/tables/ParticipantChar/Characteristics_Male_byTMAOQuantile_211111.csv")

table_save.allq <- print(table_tmao.allq, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.allq, file = "../../Outputs/tables/ParticipantChar/Characteristics_All_byTMAOQuantile_211111.csv")
```

Make similar tables but for TMAO tertile

```{r Participant characterisitics by TMAO tertile}
# TMAO table by tmao_tertile - females
factorVars <- c("age_cat", "tmao_tertile")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al", "hdl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.ft <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = phen[phen$sex == "Female",],
                                    factorVars = factorVars)
table_tmao.ft

# TMAO table by tmao_tertile - males
factorVars <- c("age_cat", "tmao_tertile")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al", "hdl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.mt <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = phen[phen$sex == "Male",], # note 6103 and 7098 have high il6
                                    factorVars = factorVars)
table_tmao.mt

# TMAO table by tmao_tertile - all, add sex
factorVars <- c("tmao_tertile")
vars <- c("age", "sex", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al", "hdl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.allt <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = phen,
                                    factorVars = factorVars)
table_tmao.allt

# Do BH correction
allt_ContTable <- attr(table_tmao.allt$ContTable, "pValues")
n <- length(allt_ContTable$pNormal)
# BH correction
allt_ContTable$padj <- stats::p.adjust(allt_ContTable$pNormal, method = "BH", n = n) 
allt_ContTable
```

Save

```{r Save characteristic tertile tables}
table_save.ft <- print(table_tmao.ft, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.ft, file = "../../Outputs/tables/ParticipantChar/Characteristics_Female_byTMAOTertile.csv")

table_save.mt <- print(table_tmao.mt, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.mt, file = "../../Outputs/tables/ParticipantChar/Characteristics_Male_byTMAOTertile.csv")

table_save.allt <- print(table_tmao.allt, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.allt, file = "../../Outputs/tables/ParticipantChar/Characteristics_All_byTMAOTertile.csv")

# BH corrections for the all tertile table
write.csv(allt_ContTable, file = "../../Outputs/tables/ParticipantChar/Characteristics_All_byTMAOTertile_Padj.csv")
```

# Cardiometabolic Risk Analysis

Note, we will remove ethnicity in these regressions because when we specifically asked if ethnicity was related to the cardiometabolic risk variables there were all non-significant relationships (uncomment the line to demonstrate for yourself). We will keep cystatin C in the regression using the same reasoning - many significant relationships.

```{r Cardiometabolic Multiple Linear Regression}
# Transfored variable list
transformed_keep_cardio_tbl <- c("sex", "age","bmi_final", "ethnicity", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al75", "hdl_bd1", "ldl_bd1", "chol_bd1", "tg_bd1_ln", "glc_bd1_ln","tnfa_bd1_ln", "il6_bd1_ln", "crp_bd1_ln", "cystatinc_bd1", "choline", "betaine", "carnitine","tmao_log")

#~~~~~~~~
# TMAO  
#~~~~~~~~
# Make a smaller df from master (the old data)
phen_sub <- phen[, colnames(phen) %in% transformed_keep_cardio_tbl]
cvd_biomark <- subset(phen_sub, select = -c(sex, age, ethnicity, cystatinc_bd1, tmao_log)) # change variable here
n <- ncol(cvd_biomark)

# LCMS method
# Run linear model
my_lm <- lapply(1:n, function(x) lm(cvd_biomark[,x] ~ tmao_log + sex + age + cystatinc_bd1, phen_sub )) # include kcal
#my_lm <- lapply(1:n, function(x) lm(cvd_biomark[,x] ~ ethnicity, phen_sub ))
#my_lm <- lapply(1:n, function(x) lm(cvd_biomark[,x] ~ cystatinc_bd1, phen_sub ))

# Organize results
my_lm2 <- sapply(my_lm, function(x){summary(x)$adj.r.squared})
my_lm2.5 <- sapply(my_lm, function(x){summary(x)$coefficients[2,1]}) # [CVD row, estimate] corresponds to CVD B 
my_lm2.6 <- sapply(my_lm, function(x){summary(x)$coefficients[2,2]}) # [CVD row, Std.Error] corresponds to CVD Std. Error 
my_lm2.7 <- sapply(my_lm, function(x){summary(x)$coefficients[2,3]}) # [CVD row, T value] corresponds to CVD t value
my_lm3 <- sapply(my_lm, function(x){summary(x)$coefficients[2,4]}) # [CVD row, P] corresponds to CVD P value
my_lm361_lcms_cvd <- cbind(my_lm2, my_lm2.5, my_lm2.6, my_lm2.7, my_lm3)
colnames(my_lm361_lcms_cvd) <- c("adj.r.squared", "estimate", "Std.Error", "T.value","P")
row.names(my_lm361_lcms_cvd) <- colnames(cvd_biomark)
res <- as.data.frame(my_lm361_lcms_cvd)

# BH correction
res$p.adj <- stats::p.adjust(res$P, method = "BH", n = n) 
```

Save table
```{r Save TMAO linear regression table}
# Save table
write.csv(res, "../../Outputs/tables/CardiometabolicMarkers/MultipleLinReg_TMAO_LCMS_n361_CVD_211111.csv")
```

## Females 
```{r Cardiometabolic Multiple Linear Regression - Females}
# Transfored variable list
transformed_keep_cardio_tbl <- c("sex", "age","bmi_final", "ethnicity", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al75", "hdl_bd1", "ldl_bd1", "chol_bd1", "tg_bd1_ln", "glc_bd1_ln","tnfa_bd1_ln", "il6_bd1_ln", "crp_bd1_ln", "cystatinc_bd1", "choline", "betaine", "carnitine","tmao_log")

#~~~~~~~~
# TMAO  
#~~~~~~~~
# Make a smaller df from master (the old data)
phen_sub <- phen[, colnames(phen) %in% transformed_keep_cardio_tbl]
# Sex data frame
phen_sub.f <- phen_sub[phen_sub$sex == "Female",]
cvd_biomark <- subset(phen_sub.f, select = -c(sex, age, ethnicity, cystatinc_bd1, tmao_log)) # change variable here
n <- ncol(cvd_biomark)

# LCMS method
# Run linear model
my_lm <- lapply(1:n, function(x) lm(cvd_biomark[,x] ~ tmao_log + age + cystatinc_bd1, phen_sub.f )) # take out sex
#my_lm <- lapply(1:n, function(x) lm(cvd_biomark[,x] ~ ethnicity, phen_sub ))
#my_lm <- lapply(1:n, function(x) lm(cvd_biomark[,x] ~ cystatinc_bd1, phen_sub ))

# Organize results
my_lm2 <- sapply(my_lm, function(x){summary(x)$adj.r.squared})
my_lm2.5 <- sapply(my_lm, function(x){summary(x)$coefficients[2,1]}) # [CVD row, estimate] corresponds to CVD B 
my_lm2.6 <- sapply(my_lm, function(x){summary(x)$coefficients[2,2]}) # [CVD row, Std.Error] corresponds to CVD Std. Error 
my_lm2.7 <- sapply(my_lm, function(x){summary(x)$coefficients[2,3]}) # [CVD row, T value] corresponds to CVD t value
my_lm3 <- sapply(my_lm, function(x){summary(x)$coefficients[2,4]}) # [CVD row, P] corresponds to CVD P value
my_lm361_lcms_cvd.f <- cbind(my_lm2, my_lm2.5, my_lm2.6, my_lm2.7, my_lm3)
colnames(my_lm361_lcms_cvd.f) <- c("adj.r.squared", "estimate", "Std.Error", "T.value","P")
row.names(my_lm361_lcms_cvd.f) <- colnames(cvd_biomark)
res.f <- as.data.frame(my_lm361_lcms_cvd.f)

# BH correction
res.f$p.adj <- stats::p.adjust(res.f$P, method = "BH", n = n) 
```

Save table
```{r Save TMAO linear regression table - Females}
# Save table
write.csv(res.f, "../../Outputs/tables/CardiometabolicMarkers/MultipleLinReg_TMAO_LCMS_n361_CVD_females_211112.csv")
```

No significant effects found in female only cohort.

## Males 
```{r Cardiometabolic Multiple Linear Regression - Males}
# Transfored variable list
transformed_keep_cardio_tbl <- c("sex", "age","bmi_final", "ethnicity", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al75", "hdl_bd1", "ldl_bd1", "chol_bd1", "tg_bd1_ln", "glc_bd1_ln","tnfa_bd1_ln", "il6_bd1_ln", "crp_bd1_ln", "cystatinc_bd1", "choline", "betaine", "carnitine","tmao_log")

#~~~~~~~~
# TMAO  
#~~~~~~~~
# Make a smaller df from master (the old data)
phen_sub <- phen[, colnames(phen) %in% transformed_keep_cardio_tbl]
# Sex data frame
phen_sub.m <- phen_sub[phen_sub$sex == "Male",]
cvd_biomark <- subset(phen_sub.m, select = -c(sex, age, ethnicity, cystatinc_bd1, tmao_log)) # change variable here
n <- ncol(cvd_biomark)

# LCMS method
# Run linear model
my_lm <- lapply(1:n, function(x) lm(cvd_biomark[,x] ~ tmao_log + age + cystatinc_bd1, phen_sub.m )) # take out sex
#my_lm <- lapply(1:n, function(x) lm(cvd_biomark[,x] ~ ethnicity, phen_sub ))
#my_lm <- lapply(1:n, function(x) lm(cvd_biomark[,x] ~ cystatinc_bd1, phen_sub ))

# Organize results
my_lm2 <- sapply(my_lm, function(x){summary(x)$adj.r.squared})
my_lm2.5 <- sapply(my_lm, function(x){summary(x)$coefficients[2,1]}) # [CVD row, estimate] corresponds to CVD B 
my_lm2.6 <- sapply(my_lm, function(x){summary(x)$coefficients[2,2]}) # [CVD row, Std.Error] corresponds to CVD Std. Error 
my_lm2.7 <- sapply(my_lm, function(x){summary(x)$coefficients[2,3]}) # [CVD row, T value] corresponds to CVD t value
my_lm3 <- sapply(my_lm, function(x){summary(x)$coefficients[2,4]}) # [CVD row, P] corresponds to CVD P value
my_lm361_lcms_cvd.m <- cbind(my_lm2, my_lm2.5, my_lm2.6, my_lm2.7, my_lm3)
colnames(my_lm361_lcms_cvd.m) <- c("adj.r.squared", "estimate", "Std.Error", "T.value","P")
row.names(my_lm361_lcms_cvd.m) <- colnames(cvd_biomark)
res.m <- as.data.frame(my_lm361_lcms_cvd.m)

# BH correction
res.m$p.adj <- stats::p.adjust(res.m$P, method = "BH", n = n) 
```

Save table
```{r Save TMAO linear regression table - Males}
# Save table
write.csv(res.m, "../../Outputs/tables/CardiometabolicMarkers/MultipleLinReg_TMAO_LCMS_n361_CVD_males_211112.csv")
```

No significant effects found in male only cohort.

# Diet Analysis

Load the diet data.

```{r Load ASA24 averaged data}
food <- readRDS(file = "../../data/processed/asa24/ASA24_meanIntake_withTransformations_211111.rds")

# Combine food with phenotype data
full <- merge(phen, food, by = 0)
row.names(full) <- full[,"Row.names"]
full <- subset(full, select = -c(Row.names))

# Make ASA24 choline log variable
full$choln_tnfs_log <- log(full$choln_tnfs)
```

Make a list of foods and covariate variables to keep.

```{r Foods to keep, results='hide'}
transformed_keep_tbl <- c("sex", 
#"sex_numeric",
"kcal_tnfs", 
"age", 
"cystatinc_bd1", 
"ethnicity",
"tmao", 
"tmao_log", 
#"TMAO.0",
#"TMAO.30",
#"TMAO.180",
#"TMAO.360",
#"lnTMAO.0",
#"lnTMAO.30",
#"lnTMAO.180",
#"lnTMAO.360",
#"TMAO_0_30",
#"TMAO_0_180",
#"TMAO_0_360",
#"Ind.SO4",
#"p.Cresol.SO4",
#"TMAO_quart_num",
#"T.pf_mps_totals_sqrRt",
"T.pf_meat_totals_sqrRt",
"T.pf_curedmeat_totals_sqrRt",
"T.pf_poult_totals_sqrRt",
"T.pf_seafd_hi_totals_Binary",
"T.pf_seafd_low_totals_Binary",
"T.pf_eggs_totals_sqrRt",
"T.pf_nutsds_totals_sqrRt",
"T.pf_legumes_totals_sqrRt",
"T.pf_total_totals_natLog",
#"T.pf_soy_totals_Binary",
"T.v_drkgr_totals_sqrRt",
"T.v_redor_total_totals_sqrRt",
#"T.v_redor_tomato_totals_sqrRt",
#"T.v_redor_other_totals_sqrRt",
"T.v_starchy_total_totals_sqrRt",
#"T.v_starchy_potato_totals_sqrRt",
#"T.v_starchy_other_totals_sqrRt",
"T.v_other_totals_sqrRt",
"T.v_total_totals_sqrRt",
#"T.v_legumes_total_sqrRt",
"T.g_whole_totals_sqrRt",
"T.g_refined_totals_sqrRt",
"g_total_totals",
"T.d_total_totals_sqrRt",
"T.d_milk_totals_sqrRt",
"T.d_cheese_totals_sqrRt",
"T.d_yogurt_totals_Binary",
"choln_tnfs_log"
)
```

Now run the multiple linear regression. Note, we will not include ethnicity in these regressions because although TMAO levels significantly differ between ethnic groups, there was no relationship between TMAO and ethnicity. Additionally, diet patterns differ by ethnicity so we could be over adjusting.

```{r TMAO ~ food - LCMS BD1 - lm 356 ppl, results='hide'}
# Make a smaller df from master (the old data)
full_sub <- full[, colnames(full) %in% transformed_keep_tbl]
food <- subset(full_sub, select = -c(sex, kcal_tnfs, age, cystatinc_bd1, ethnicity, tmao, tmao_log))
n <- ncol(food)

# LCMS method
# Run linear model
my_lm <- lapply(1:n, function(x) lm(tmao_log ~ food[,x] + sex*age + cystatinc_bd1 + kcal_tnfs, full_sub )) # include kcal

# Organize results
my_lm2 <- sapply(my_lm, function(x){summary(x)$adj.r.squared})
my_lm2.5 <- sapply(my_lm, function(x){summary(x)$coefficients[2,1]}) # [food row, estimate] corresponds to food B 
my_lm2.6 <- sapply(my_lm, function(x){summary(x)$coefficients[2,2]}) # [food row, Std.Error] corresponds to food Std. Error 
my_lm2.7 <- sapply(my_lm, function(x){summary(x)$coefficients[2,3]}) # [food row, T value] corresponds to food t value
my_lm3 <- sapply(my_lm, function(x){summary(x)$coefficients[2,4]}) # [food row, P] corresponds to food P value
my_lm356_lcms <- cbind(my_lm2, my_lm2.5, my_lm2.6, my_lm2.7, my_lm3)
colnames(my_lm356_lcms) <- c("adj.r.squared", "estimate", "Std.Error", "T.value","P")
row.names(my_lm356_lcms) <- colnames(food)
res_food <- as.data.frame(my_lm356_lcms)

# BH correction
res_food$p.adj <- stats::p.adjust(res_food$P, method = "BH", n = n) 
```

Save
```{r save food linear regression table}
# Save table
write.csv(res_food, "../../Outputs/tables/Diet/MultipleLinReg_TMAO_LCMS_n356_ASA24_211111.csv")
```

Make TMAO food intake by quantile table
```{r Food Intake by TMAO quantile}
# TMAO table by age bin - females
factorVars <- c("age_cat", "tmao_quantile")
vars <- c("pf_meat_totals", 
          "pf_curedmeat_totals",
          "pf_poult_totals",
          "pf_seafd_hi_totals",
          "pf_seafd_low_totals",
          "pf_eggs_totals",
          "pf_nutsds_totals",
          "pf_legumes_totals",
          "pf_total_totals",
          #"pf_soy_totals_Binary",
          "v_drkgr_totals",
          "v_redor_total_totals",
          #"v_redor_tomato_totals",
          #"v_redor_other_totals",
          "v_starchy_total_totals",
          #"v_starchy_potato_totals",
          #"v_starchy_other_totals",
          "v_other_totals",
          "v_total_totals",
          #"v_legumes_total",
          "g_whole_totals",
          "g_refined_totals",
          "g_total_totals",
          "d_total_totals",
          "d_milk_totals",
          "d_cheese_totals",
          "d_yogurt_totals",
          "hei_asa24_totalscore",
          "hei_asa24_totalveg",
          "hei_asa24_sfat",
          "hei_asa24_",
          "choln_tnfs",
          "age",
          "ethnicity" ,
          "bmi_final",
          "tmao",
          "choline",
          "betaine",
          "carnitine")
table_food.fq <- CreateTableOne(vars = vars,
                                    strata = "tmao_quantile",
                                    data = full[full$sex == "Female",],
                                    factorVars = factorVars)
table_food.fq

# TMAO table by age bin - males
table_food.mq <- CreateTableOne(vars = vars,
                                    strata = "tmao_quantile",
                                    data = full[full$sex == "Male",],
                                    factorVars = factorVars)
table_food.mq

# TMAO table by sex
table_food.allq <- CreateTableOne(vars = vars,
                                    strata = "tmao_quantile",
                                    data = full,
                                    factorVars = factorVars)
table_food.allq
```

```{r Save food quantile tables}
table_save.fq <- print(table_food.fq, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.fq, file = "../../Outputs/tables/ParticipantChar/Food_Female_byTMAOQuantile_211111.csv")

table_save.mq <- print(table_food.mq, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.mq, file = "../../Outputs/tables/ParticipantChar/Food_Male_byTMAOQuantile_211111.csv")

table_save.allq <- print(table_food.allq, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.allq, file = "../../Outputs/tables/ParticipantChar/Food_All_byTMAOQuantile_211111.csv")
```

TMAO tertile 
Make TMAO food intake by tertile table
```{r Food Intake by TMAO tertile}
# TMAO table by age bin - females
factorVars <- c("age_cat", "tmao_quantile")
vars <- c("pf_meat_totals", 
          "pf_curedmeat_totals",
          "pf_poult_totals",
          "pf_seafd_hi_totals",
          "pf_seafd_low_totals",
          "pf_eggs_totals",
          "pf_nutsds_totals",
          "pf_legumes_totals",
          "pf_total_totals",
          #"pf_soy_totals_Binary",
          "v_drkgr_totals",
          "v_redor_total_totals",
          #"v_redor_tomato_totals",
          #"v_redor_other_totals",
          "v_starchy_total_totals",
          #"v_starchy_potato_totals",
          #"v_starchy_other_totals",
          "v_other_totals",
          "v_total_totals",
          #"v_legumes_total",
          "g_whole_totals",
          "g_refined_totals",
          "g_total_totals",
          "d_total_totals",
          "d_milk_totals",
          "d_cheese_totals",
          "d_yogurt_totals",
          "hei_asa24_totalscore",
          "hei_asa24_totalveg",
          "hei_asa24_sfat",
          "hei_asa24_",
          "choln_tnfs",
          "age",
          "ethnicity" ,
          "bmi_final",
          "tmao",
          "choline",
          "betaine",
          "carnitine")
table_food.ft <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = full[full$sex == "Female",],
                                    factorVars = factorVars)
table_food.ft

# TMAO table by age bin - males
table_food.mt <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = full[full$sex == "Male",],
                                    factorVars = factorVars)
table_food.mt

# TMAO table by sex
table_food.allt <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = full,
                                    factorVars = factorVars)
table_food.allt

# Do BH correction
food_allt_ContTable <- attr(table_food.allt$ContTable, "pValues")
n <- length(food_allt_ContTable$pNormal)
# BH correction
food_allt_ContTable$padj <- stats::p.adjust(food_allt_ContTable$pNormal, method = "BH", n = n) 
food_allt_ContTable
```

```{r Save food tertile tables}
table_save.ft <- print(table_food.ft, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.ft, file = "../../Outputs/tables/ParticipantChar/Food_Female_byTMAOTertile.csv")

table_save.mt <- print(table_food.mt, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.mt, file = "../../Outputs/tables/ParticipantChar/Food_Male_byTMAOTertile.csv")

table_save.allt <- print(table_food.allt, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.allt, file = "../../Outputs/tables/ParticipantChar/Food_All_byTMAOTertile.csv")

write.csv(food_allt_ContTable, file = "../../Outputs/tables/ParticipantChar/Food_All_byTMAOTertile_PadjTable.csv")
```

# Miscellaneous for paper

Demonstrate the sex:age interaction for females.
```{r sex*age interaction}
summary(lm(tmao_log ~ sex*age + ethnicity + cystatinc_bd1, phen))
```

Describe meat intake in the cohort

```{r Meat intake}
# Beef, veal, pork, lamb, and game meat; excludes organ meat and cured meat (oz. eq.)
summary(full$pf_meat_totals)
sd(full$pf_meat_totals)

# Frankfurters, sausages, corned beef, and luncheon meat that are made from beef, pork, or poultry (oz. eq.)
summary(full$pf_curedmeat_totals)
sd(full$pf_curedmeat_totals)

# Chicken, turkey, Cornish hens, duck, goose, quail, and pheasant (game birds); excludes organ meat and cured meat (oz. eq.)
summary(full$pf_poult_totals)
sd(full$pf_poult_totals)

# Total meat, poultry, organ meat, cured meat, seafood, eggs, soy, and nuts and seeds; excludes legumes (oz. eq.)
summary(full$pf_total_totals)
sd(full$pf_total_totals)

# Total of meat, poultry, seafood, organ meat, and cured meat (oz. eq.)
summary(full$pf_mps_total_totals)
sd(full$pf_mps_total_totals)

# Eggs (chicken, duck, goose, quail) and egg substitutes (oz. eq.)
summary(full$pf_eggs_totals)
sd(full$pf_eggs_totals)

# HEI total score
summary(full$hei_asa24_totalscore)
sd(full$hei_asa24_totalscore, na.rm = TRUE)

# Kcal
summary(full$kcal_tnfs)
sd(full$kcal_tnfs)
```

"waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al75", "hdl_bd1", "ldl_bd1", "chol_bd1", "tg_bd1_ln", "glc_bd1_ln","tnfa_bd1_ln", "il6_bd1_ln", "crp_bd1_ln",

BS section below...
```{r}
full$low_mps_meat <- ifelse(full$pf_mps_total_totals <= median(full$pf_mps_total_totals), "low", "high")
full$low_mps_meat <- ifelse(full$pf_mps_total_totals <= 9.091, "low", "high")
summary(lm(glc_bd1_ln ~  tmao_log + sex*age + cystatinc_bd1, full[full$low_mps_meat == "high",]))
summary(lm(glc_bd1_ln ~  tmao_log + sex*age + cystatinc_bd1, full[full$low_mps_meat == "high",]))
```