### 2022-03-12 Rmarkdown/analysis/Response_to_Reviewers.Rmd
Response to reviewer comments

---
title: "Response to Reviewers"
author: "Kristen James"
date: "3/7/2022"
output: html_document
---

This markdown was created in response to the Reviewer's comments during the peer-review process of our manuscript submitted to Nutrients. The workflow is borrowed in large part from "Characteristics_Diet_CardiometabolicRisk_211111.Rmd" with subtle changes.

Load libraries

```{r Libraries, warning=FALSE, message=FALSE}
library(plyr)
library(tableone)
```

Load data

```{r Load data}
phen <- readRDS("../../data/processed/phenotype/Phenotypes_211108.rds")

# Complete TMAO data only
phen <- phen[complete.cases(phen$tmao == TRUE),]

# Get rid of NA ethnicity category
phen$ethnicity <- droplevels(phen$ethnicity)

# Factor and recode sex
phen$sex <- as.factor(phen$sex)
phen$sex <- plyr::revalue(phen$sex, c("1" = "Male", "2" = "Female"))

# Make TMAO log variable
phen$tmao_log <- log(phen$tmao)

```

# Reviewer 1

Provide the number of participants in each age-sex-BMI bin to answer Reviewer 1's comment about statistical power.

```{r}
table(phen$bin_number)
```

Reviewer 1 didn't understand that the relationship between TMAO and age differed by sex. We will make it more clear by stating the restults of the statistical test.

```{r sex*age interaction}
summary(lm(tmao_log ~ sex*age  + cystatinc_bd1, phen))
```

# Reviewer 2

## Fiber Comment
Reviewer 2 asked why we did not include fiber in the analysis. We will add total fiber intake from foods and supplements (fibe_tnfs) to our analysis.

First, lets set up the dataframe. This code is borrowed from "Characteristics_Diet_CardiometabolicRisk_211111.Rmd".

```{r Load data and preprocess}
phen <- readRDS("../../data/processed/phenotype/Phenotypes_211108.rds")

# Complete TMAO data only
phen <- phen[complete.cases(phen$tmao == TRUE),]

# Get rid of NA ethnicity category
phen$ethnicity <- droplevels(phen$ethnicity)

# Factor and recode sex
phen$sex <- as.factor(phen$sex)
phen$sex <- plyr::revalue(phen$sex, c("1" = "Male", "2" = "Female"))

# Make TMAO log variable
phen$tmao_log <- log(phen$tmao)

# Convert crp units
phen$crp_bd1.mgdL <- phen$crp_bd1 * 0.0001
```

Create quantile and tertile classifiers. Transform data to comply to normality (Shapiro-Wilk test W statistic > 0.95).

```{r Quantile and Tertile TMAO}
# Quantile
tmao_quantile <- quantile(phen$tmao)
phen$tmao_quantile <- ifelse(phen$tmao <= tmao_quantile[2], "Quantile1",
                             ifelse(phen$tmao > tmao_quantile[2] & phen$tmao <= tmao_quantile[3], "Quantile2",
                                   ifelse(phen$tmao > tmao_quantile[3] & phen$tmao <= tmao_quantile[4], "Quantile3", "Quantile4")))

# Check it
table(phen$tmao_quantile)
phen$tmao_quantile <- as.factor(phen$tmao_quantile)
str(phen$tmao_quantile)

# Tertile
tmao_tertile <- quantile(phen$tmao, c(0:3/3))
phen$tmao_tertile <- ifelse(phen$tmao <= tmao_tertile[2], "Tertile1",
                             ifelse(phen$tmao > tmao_tertile[2] & phen$tmao <= tmao_tertile[3], "Tertile2", "Tertile3"))
# Factor it
phen$tmao_tertile <- factor(phen$tmao_tertile)
table(phen$tmao_tertile)

# Print tertiles for manuscript
tmao_tertile

# Checked all variables for normality via shapiro.test() in console. TG, and glc ere the only ones that didn't pass.
phen$glc_bd1_ln <- log(phen$glc_bd1)
phen$tg_bd1_ln <- log(phen$tg_bd1)
phen$crp_bd1_ln <- log(phen$crp_bd1)
phen$tnfa_bd1_ln <- log(phen$tnfa_bd1)
phen$il6_bd1_ln <- log(phen$il6_bd1)
phen$insulin_bd1_ln <- log(phen$insulin_bd1)
```

Load the diet data.

```{r Load ASA24 averaged data}
food <- readRDS(file = "../../data/processed/asa24/ASA24_meanIntake_withTransformations_211111.rds")

# Combine food with phenotype data
full <- merge(phen, food, by = 0)
row.names(full) <- full[,"Row.names"]
full <- subset(full, select = -c(Row.names))

# Make ASA24 choline log variable
full$choln_tnfs_log <- log(full$choln_tnfs)

# Make ASA24 fiber log variable
shapiro.test(log(full$fibe_tnfs))
full$fibe_tnfs_log <- log(full$fibe_tnfs)
```

Make a list of foods and covariate variables to keep. Add fiber this time!

```{r Foods to keep, results='hide'}
transformed_keep_tbl <- c("sex", 
#"sex_numeric",
"kcal_tnfs", 
"age", 
"cystatinc_bd1", 
"ethnicity",
"tmao", 
"tmao_log", 
#"TMAO.0",
#"TMAO.30",
#"TMAO.180",
#"TMAO.360",
#"lnTMAO.0",
#"lnTMAO.30",
#"lnTMAO.180",
#"lnTMAO.360",
#"TMAO_0_30",
#"TMAO_0_180",
#"TMAO_0_360",
#"Ind.SO4",
#"p.Cresol.SO4",
#"TMAO_quart_num",
#"T.pf_mps_totals_sqrRt",
"T.pf_meat_totals_sqrRt",
"T.pf_curedmeat_totals_sqrRt",
"T.pf_poult_totals_sqrRt",
"T.pf_seafd_hi_totals_Binary",
"T.pf_seafd_low_totals_Binary",
"T.pf_eggs_totals_sqrRt",
"T.pf_nutsds_totals_sqrRt",
"T.pf_legumes_totals_sqrRt",
"T.pf_total_totals_natLog",
#"T.pf_soy_totals_Binary",
"T.v_drkgr_totals_sqrRt",
"T.v_redor_total_totals_sqrRt",
#"T.v_redor_tomato_totals_sqrRt",
#"T.v_redor_other_totals_sqrRt",
"T.v_starchy_total_totals_sqrRt",
#"T.v_starchy_potato_totals_sqrRt",
#"T.v_starchy_other_totals_sqrRt",
"T.v_other_totals_sqrRt",
"T.v_total_totals_sqrRt",
#"T.v_legumes_total_sqrRt",
"T.g_whole_totals_sqrRt",
"T.g_refined_totals_sqrRt",
"g_total_totals",
"T.d_total_totals_sqrRt",
"T.d_milk_totals_sqrRt",
"T.d_cheese_totals_sqrRt",
"T.d_yogurt_totals_Binary",
"choln_tnfs_log",
"hei_asa24_totalscore",
"fibe_tnfs_log"
)
```

Now run the multiple linear regression. Note, we will not include ethnicity in these regressions because although TMAO levels significantly differ between ethnic groups, there was no relationship between TMAO and ethnicity. Additionally, diet patterns differ by ethnicity so we could be over adjusting.

```{r TMAO ~ food - LCMS BD1 - lm 356 ppl, results='hide'}
# Make a smaller df from master (the old data)
full_sub <- full[, colnames(full) %in% transformed_keep_tbl]
food <- subset(full_sub, select = -c(sex, kcal_tnfs, age, cystatinc_bd1, ethnicity, tmao, tmao_log))
n <- ncol(food)

# LCMS method
# Run linear model
my_lm <- lapply(1:n, function(x) lm(tmao_log ~ food[,x] + sex*age + cystatinc_bd1 + kcal_tnfs, full_sub )) # include kcal

# Organize results
my_lm2 <- sapply(my_lm, function(x){summary(x)$adj.r.squared})
my_lm2.5 <- sapply(my_lm, function(x){summary(x)$coefficients[2,1]}) # [food row, estimate] corresponds to food B 
my_lm2.6 <- sapply(my_lm, function(x){summary(x)$coefficients[2,2]}) # [food row, Std.Error] corresponds to food Std. Error 
my_lm2.7 <- sapply(my_lm, function(x){summary(x)$coefficients[2,3]}) # [food row, T value] corresponds to food t value
my_lm3 <- sapply(my_lm, function(x){summary(x)$coefficients[2,4]}) # [food row, P] corresponds to food P value
my_lm356_lcms <- cbind(my_lm2, my_lm2.5, my_lm2.6, my_lm2.7, my_lm3)
colnames(my_lm356_lcms) <- c("adj.r.squared", "estimate", "Std.Error", "T.value","P")
row.names(my_lm356_lcms) <- colnames(food)
res_food <- as.data.frame(my_lm356_lcms)

# BH correction
res_food$p.adj <- stats::p.adjust(res_food$P, method = "BH", n = n) 
```

Save with the extension "_withFiber"

```{r save food linear regression table}
# Save table
write.csv(res_food, "../../Outputs/tables/Diet/MultipleLinReg_TMAO_LCMS_n356_ASA24_withFiber.csv")
```

We also need to add fiber to our food by TMAO tertile table.

```{r Food Intake by TMAO tertile}
# TMAO table by age bin - females
factorVars <- c("age_cat", "tmao_quantile")
vars <- c("pf_meat_totals", 
          "pf_curedmeat_totals",
          "pf_poult_totals",
          "pf_seafd_hi_totals",
          "pf_seafd_low_totals",
          "pf_eggs_totals",
          "pf_nutsds_totals",
          "pf_legumes_totals",
          "pf_total_totals",
          #"pf_soy_totals_Binary",
          "v_drkgr_totals",
          "v_redor_total_totals",
          #"v_redor_tomato_totals",
          #"v_redor_other_totals",
          "v_starchy_total_totals",
          #"v_starchy_potato_totals",
          #"v_starchy_other_totals",
          "v_other_totals",
          "v_total_totals",
          #"v_legumes_total",
          "g_whole_totals",
          "g_refined_totals",
          "g_total_totals",
          "d_total_totals",
          "d_milk_totals",
          "d_cheese_totals",
          "d_yogurt_totals",
          "hei_asa24_totalscore",
          "choln_tnfs",
          "fibe_tnfs", # fiber added!
          "age",
          "ethnicity" ,
          "bmi_final",
          "tmao",
          "choline",
          "betaine",
          "carnitine")
table_food.ft <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = full[full$sex == "Female",],
                                    factorVars = factorVars)
table_food.ft

# TMAO table by age bin - males
table_food.mt <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = full[full$sex == "Male",],
                                    factorVars = factorVars)
table_food.mt

# TMAO table by sex
table_food.allt <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = full,
                                    factorVars = factorVars)
table_food.allt

# Do BH correction
food_allt_ContTable <- attr(table_food.allt$ContTable, "pValues")
n <- length(food_allt_ContTable$pNormal)
# BH correction
food_allt_ContTable$padj <- stats::p.adjust(food_allt_ContTable$pNormal, method = "BH", n = n) 
food_allt_ContTable
```

Save with the "_withFiber" extension.

```{r Save food tertile tables}
table_save.ft <- print(table_food.ft, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.ft, file = "../../Outputs/tables/ParticipantChar/Food_Female_byTMAOTertile_withFiber.csv")

table_save.mt <- print(table_food.mt, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.mt, file = "../../Outputs/tables/ParticipantChar/Food_Male_byTMAOTertile_withFiber.csv")

table_save.allt <- print(table_food.allt, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.allt, file = "../../Outputs/tables/ParticipantChar/Food_All_byTMAOTertile_withFiber.csv")

write.csv(food_allt_ContTable, file = "../../Outputs/tables/ParticipantChar/Food_All_byTMAOTertile_PadjTable_withFiber.csv")
```

Reviewer 2 asked us to clarify what transformations were used in the endothelial analysis. Ultimately, we used AI in Table 1 and  AI75 in the linear regression analysis. This was because when we use AI75 in the descriptive table, the mean values are negative and positive, which is confusing/not intuitive. The following chunk demonstrates this point.

```{r Participant characterisitics by TMAO tertile}
# TMAO table by tmao_tertile - females
factorVars <- c("age_cat", "tmao_tertile")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al75", "hdl_bd1", "ldl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "insulin_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.ft <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = phen[phen$sex == "Female",],
                                    factorVars = factorVars)
table_tmao.ft

# TMAO table by tmao_tertile - males
factorVars <- c("age_cat", "tmao_tertile")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al75", "hdl_bd1", "ldl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "insulin_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.mt <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = phen[phen$sex == "Male",], # note 6103 and 7098 have high il6
                                    factorVars = factorVars)
table_tmao.mt

# TMAO table by tmao_tertile - all, add sex
factorVars <- c("tmao_tertile", "sex")
vars <- c("age", "sex", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al75", "hdl_bd1", "ldl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "insulin_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.allt <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = phen,
                                    factorVars = factorVars)
table_tmao.allt

# Do BH correction for continuous variables
allt_ContTable <- attr(table_tmao.allt$ContTable, "pValues")
n <- length(allt_ContTable$pNormal)
# BH correction
allt_ContTable$padj <- stats::p.adjust(allt_ContTable$pNormal, method = "BH", n = n) 
allt_ContTable

# Do BH correction for categorical variables
allt_CatTable <- attr(table_tmao.allt$CatTable, "pValues")
n <- 23
# BH correction
allt_CatTable$padj <- stats::p.adjust(allt_CatTable$pApprox, method = "BH", n = n) 
allt_CatTable
```

Save with "_al75" extension; however, we did not use these tables (we used the AI version originally created).

```{r Save characteristic tertile tables}
table_save.ft <- print(table_tmao.ft, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.ft, file = "../../Outputs/tables/ParticipantChar/Characteristics_Female_byTMAOTertile_al75.csv")

table_save.mt <- print(table_tmao.mt, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.mt, file = "../../Outputs/tables/ParticipantChar/Characteristics_Male_byTMAOTertile_al75.csv")

table_save.allt <- print(table_tmao.allt, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.allt, file = "../../Outputs/tables/ParticipantChar/Characteristics_All_byTMAOTertile_al75.csv")

# BH corrections for the all tertile table
write.csv(allt_ContTable, file = "../../Outputs/tables/ParticipantChar/Characteristics_All_byTMAOTertile_Padj_al75.csv")
```

Do some individuals have undiagnosed hyperglycemia or hyperlipidemia? Yes

```{r undiagnosed metabolic disease}
summary(phen$glc_bd1)
summary(phen$tg_bd1)
summary(phen$chol_bd1)

hist(phen$glc_bd1, breaks = 40)
hist(phen$tg_bd1, breaks = 40)
hist(phen$chol_bd1, breaks = 40)
```